### 2024-09-21 paper/paper.qmd
Update analysis, data cleaning, and testing scripts, and regenerate paper

---
title: "Analysis of Public Opinion Polls Conducted by the City of Toronto"
subtitle: "Insights on Front Yard Parking, Traffic Calming, and Boulevard Cafes"
author:
  - "Tim Chen"
thanks: "Code and data are available at: https://github.com/timchen0326/sta304_paper_1."
date: today
date-format: long
abstract: "This paper analyzes data from the City of Toronto's public opinion polls on key municipal projects. We use a dataset covering several initiatives, including front yard parking, boulevard cafes, and traffic calming. Our analysis reveals significant variation in voter turnout and approval rates across these different projects. The findings provide insights into community engagement and policy preferences, with potential implications for future urban planning and public consultation processes"
format: 
  pdf:
    toc: true        # Enables Table of Contents
    toc-depth: 2     # Specifies depth level of headings included in the ToC
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
# Install and load the required packages
options(repos = c(CRAN = "https://cran.rstudio.com/"))

data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
```

# Introduction

Public opinion polls are an essential part of local governance, offering insights into how communities perceive proposed municipal projects. In the City of Toronto, polls are regularly conducted on topics such as front yard parking, traffic calming, and the installation of boulevard cafes. These polls not only influence policy decisions but also serve as a proxy for measuring community engagement.

The purpose of this paper is to explore the data collected from these polls and to uncover key trends in public opinion. Specifically, we analyze voter turnout, the proportion of favorable outcomes, and how these vary across different types of projects.

The remainder of this paper is structured as follows. @sec-data provides a detailed description of the dataset used. @sec-results presents the analysis of the data, including graphical representations of the results. Finally, @sec-dis discusses the findings and concludes with potential implications for future polling processes.

# Data {#sec-data}

The dataset used for this analysis was sourced from @opendatatorotno and conducted using @citeR. It contains information on several public polls, including fields such as the number of ballots cast, ballots in favor, ballots opposed, and poll results. The dataset spans a period from 2015 to 2023 and covers various municipal initiatives.

The following figure shows the distribution of the number of ballots cast for different types of municipal applications, providing an overview of public participation.
```{r}
#| label: fig-ballots-cast
#| fig-cap: Distribution of Ballots Cast by Application Type
#| echo: false

data %>%
  ggplot(aes(x = application_for, y = ballots_cast, fill = application_for)) +
  geom_boxplot() +
  labs(x = "Application Type", y = "Number of Ballots Cast", title = "Distribution of Ballots Cast by Application Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

@tbl-poll-results provides a summary of the poll results for different application types, including the number of "In Favour" and "Opposed" results for each type.

```{r}
#| label: tbl-poll-results
#| echo: false
#| tbl-cap: "Summary of Poll Results"
data %>%
  group_by(application_for, poll_result) %>%
  summarise(count = n(), .groups = 'drop') %>%
  knitr::kable(col.names = c("Application Type", "Poll Result", "Count"))

```

The dataset also contains information on whether the response rate was met, which is crucial for validating the results of a poll. The following figure shows the proportion of polls that met the response rate requirement.

```{r}
#| label: fig-response-rate
#| fig-cap: Response Rate Met by Application Type
#| echo: false

data %>%
  ggplot(aes(x = application_for, fill = response_rate_met)) +
  geom_bar(position = "dodge") +
  labs(x = "Application Type", y = "Number of Polls", fill = "Response Rate Met", title = "Response Rate Met by Application Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

# Results {#sec-results}

## Voter Turnout
Voter turnout, as measured by the number of ballots cast, varies significantly across different application types. For example, Traffic calming initiatives attracted an average of 200 ballots per poll, while front yard parking attracted an average of 100 ballots, as seen in @fig-ballots-cast, while front yard parking proposals generally receive fewer votes. This suggests that residents are more engaged with policies that directly affect public infrastructure and safety.

## Poll Results
The majority of polls analyzed resulted in a favorable outcome ("In Favour"), particularly for front yard parking and boulevard cafes. As shown in , most applications were supported by the public, although there were notable exceptions in the case of traffic calming initiatives, which received a higher number of "Opposed" votes.

## Response Rate
Meeting the response rate requirement is essential for validating poll results. As shown in @fig-response-rate, response rates were more likely to be met for boulevard cafes and front yard parking polls. This could indicate a higher level of engagement in these areas, possibly due to their direct impact on property use.

# Discussion {#sec-dis}

## Key Findings
One of the most significant findings in this analysis is the variation in voter turnout based on the type of municipal application. Traffic calming initiatives, which have a direct impact on road use and safety, attracted the highest number of ballots cast, indicating strong public interest in infrastructure-related projects.

In contrast, front yard parking polls typically received lower voter turnout, possibly because they are seen as affecting a smaller portion of the population. However, these polls still garnered high levels of support from those who participated.

## Limitations and Future Work
One limitation of this study is the potential for missing data. For instance, certain polls lack complete address information, which could impact the analysis of geographic trends in polling results. Future work could explore more granular patterns in voter engagement by incorporating additional demographic and geographic data.

Additionally, further research could investigate the long-term outcomes of these polls, examining whether the results led to sustained community satisfaction with the implemented policies.

\newpage

\appendix

# Appendix {-}

\newpage

# References